Binarity's effects on occurrence rates measured by transit surveys

### 2017-10-19 paper/main.tex
moving things around

%\documentclass[12pt,preprint]{aastex}
%\documentclass[preprint2,12pt]{aastex}
\documentclass{emulateapj}
%\usepackage{apjfonts}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{todonotes}

@@

% ------------------------------------------------------------------------
% New commands
%
\def\ltsima{$\; \buildrel < \over \sim \;$}
\def\lsim{\lower.5ex\hbox{\ltsima}}
\def\gtsima{$\; \buildrel > \over \sim \;$}
\def\gsim{\lower.5ex\hbox{\gtsima}}
\def\tess{{\it TESS} }
\def \teff {T_{\rm eff}}
\def \phir {\Phi_{\rm R}}
\def \fov {24$^{\circ}$}
\def \pixsz {21.1''}
\def \aeff {69.1 cm$^2$ }    
\def \epd {105 mm}

% -------------------------------------------------------------------------
%

\bibliographystyle{apj}

\title{ Analytic Transit Surveys }

@@

% \journalinfo{Draft version}
\slugcomment{Memo for internal use}

%\altaffiltext{1}{Princeton University}

\begin{abstract}

We describe a few different toy transit surveys.

First, we derive the signal to noise distribution of planets detected in a 
survey where all the stars and planets are the same (except some stars are twin 
binary systems). We give expressions for the number of detected planets, and 
for possible errors one can make when deriving occurrence rates.

Then, we allow the properties of the secondary in binary systems to vary. A 
distribution of light ratios introduces additional biases, which we discuss.
We reassess the errors related to occurrence rates.

\end{abstract}

\keywords{planets and satellites:\ detection}

\section{Model \#1: fixed stars, fixed planets, fixed light-ratio binaries}
\label{sec:model_1}

Ah, transiting planets! We learn so much by studying them. But how much can we 
actually learn, and how much is messed up by binarity?

Imagine the following transit survey:
\begin{enumerate}
\item You are going to observe the entire sky for a duration $T_{\rm obs}$, 
with a detector of area $A$, and known bandpass. Your detector is photon-noise 
limited.
%
\item You are interested only in detecting planets of radius $R_p$, and orbital 
period $P$. For instance, $R_p=R_\oplus$, $P={\rm 1\ year}$.
%
\item You are only interested in detecting them around stars of radius $R_1$, 
and luminosity $L_1$. For instance, G2V dwarfs.
%
\item You can only detect your `nominal planet' when you observe signals with
${\rm S/N} > {\rm (S/N)_{min}}$.
For a photon-noise limited survey, this minimum signal to noise ratio is 
equivalent to a minimum flux, $F_{\rm min}$.
%
\item However, you opt to observe all the ``points'' on-sky with apparent 
magnitude $m < m_{\rm lim}$, or equivalently with energy flux in your bandpass 
greater than some limit $F > F_{\rm lim}$.
(You will not be able to detect planets around the faintest of 
these, and will need to perform a completeness correction to later derive 
occurrence rates.)
Generally $F_{\rm lim} < F_{\rm min}$, \textit{e.g.}, whenever the original 
magnitude cut defines a ``source catalog'' from which ``searchable stars'' are 
selected, as in the KIC [Batalha et al, 2010, Brown et al 2011] and the TIC 
[Stassun et al 2017].
\end{enumerate}

You get funding, and perform the S/N-limited survey. All planets with 
${\rm S/N} > {\rm (S/N)_{min}}$ are detected. The rest are not.
You now wish to derive an occurrence rate for planets of radius $R_p$ and 
orbital period $P$.
Assume your universe is a universe in which:
\begin{itemize}
	\item The true population of ``points'' (stellar systems, all unresolved) 
	comprises both single and double star systems. Single star systems have 
	luminosity in the observed bandpass $L_1$, radii $R_1$, and effective 
	temperature $T_{\rm eff,1}$.
	Double star systems have luminosity in the observed bandpass $L_d = 
	(1+\gamma_R)L_1$, for $\gamma_R = L_2/L_1$ the ratio of the luminosity of 
	the secondary to the primary. 
	In this section, $\gamma_R$ is a constant across the population 
	of star systems.
	The ratio of the two number densities in a 
	volume-limited sample is the binary fraction\footnote{The binary fraction 
		is equivalent to the multiplicity fraction if there are no triple, 
		quadruple, $\ldots$ systems.}.
	\item The true population of planets around these stars is as follows:
	\subitem A fraction $\Gamma_{t,s}$ of stars in single star systems 
	have a planet of radius $R_p$, with orbital period $P$.
	\subitem A fraction $\Gamma_{t,d}$ of each star in a double star 
	system has a planet of radius $R_p$, with orbital period $P$. For instance, 
	if $\Gamma_{t,s} = \Gamma_{t,d} = 0.1$, on average each double 
	system contributes 0.2 planets, and each single system 0.1 planets.
	Any astrophysical difference in planet formation between singles and 
	binaries is captured by these two terms.
	Note that we are taking an limit by assuming the primary and secondary of 
	a binary system host planets at equal rates. If the secondary is 
	non-identical to the primary, it's clearly wrong.	
\end{itemize}

For simplicity, one might wish to imagine the observer correctly 
pre-selects all of the ``searchable stars'', in the style of [Pepper et al, 
2003].
In other words, they would begin by knowing ${\rm (S/N)_{min}}$, and then only 
observe stars for which they would have adequate flux to achieve it -- setting 
$F_{\rm lim} = F_{\rm min}$.
This simplification would naively allow the observer to ignore 
completeness effects when deriving occurrence rates, since their completeness 
is 100\%.

We do not begin this way because in a stellar population of both singles and 
binaries, the above approach ignores necessary incompleteness for 
binary systems.
If we set the limiting flux to give 100\% completeness for single stars, we 
will unwittingly select binaries near that flux limit, for which dilution 
will push transit signals below the detection threshold.
This is because a flux limit maps onto three different maximum 
detection distances for a population of single stars and twin binaries:
\begin{enumerate}
\item $d_{{\rm max},s}$: the maximum distance out to which single stars are
  selected. This can be chosen to coincide with the maximum distance out to
  which planets are detectable around single stars.
%
\item $d_{{\rm max},d}^\star$: the maximum distance out to which double stars
  are selected. This is different from $d_{{\rm max},d}^{\rm p}$, the maximum
  distance out to which planets are detectable around double stars. For a 
  population with fixed  $\gamma_R$, since ${\rm S/N} \propto \mathcal{D} 
  L^{-1/2} d$,
  \begin{equation}
    d_{{\rm max},d}^{\rm p} = (1 + \gamma_R)^{-1/2} d_{{\rm max},s} = 
    (1 + \gamma_R)^{-1} d_{{\rm max},d}^\star.
  \end{equation}
  For $\gamma_R = 1$, this means only 1 in 8 selected binary stars can yield
  detectable planets.
\end{enumerate}
We develop tools to address this and related completeness effects from the 
outset.

Consider the following questions:
\begin{enumerate}
\item How many single and double star systems, respectively, are in the sample?
Correspondingly, how many stars are in the sample?

\item How many planets are in the sample? (Orbiting single stars, and orbiting 
double stars respectively).

\item What is the true occurrence rate?

\item How many planets are detected?

\item What occurrence rate does astronomer A, who has never heard of binary 
star systems, or completeness, derive for planets of radius $R_p$ and period 
$P$?

\item What occurrence rate does astronomer B, who accounts for the ``2 for 1'' 
effect of binarity (\textit{i.e.} that the sample actually has more stars than 
astronomer A thought) derive? (Astronomer B neglects completeness).

\item What about astronomer C, who accounts for ``2 for 1'' \textit{and} 
misclassification due to diluted radii? In other words, astronomer C did a 
combination of high resolution imaging and RV followup on every candidate, and
correctly classifies the planetary radii in every case.

\item What about astronomer D, who additionally notes the importance of 
completeness?
\end{enumerate}

\subsection{How many stars are in the sample?}

Let $N_s$ be the number of single star systems, and $N_d$ the number of double 
star systems. Then the total number of stars in the sample is
\begin{equation}
N_{\rm stars} = N_s + 2N_d.
\end{equation}

In a magnitude-limited sample in which stars are uniformly distributed in 
volume, the number of stars will be the number density times the volume.
If the volume is taken to be a sphere over which the number density is uniform,
\begin{equation}
N_i = n_{i} \frac{4\pi}{3} d_{{\rm max},i}^3,
\label{eq:number_systems}
\end{equation}
for $i\in{\rm \{single,double\} } \equiv { \{s,d\} }$, and
\begin{equation}
\frac{n_{d}}{n_{s}} = {\rm binary\ fraction} \equiv {\rm BF}
\end{equation}
by definition. The absolute normalization of the number density is a measured 
quantity, as is the binary fraction. For G2V dwarfs, the latter is $\approx 
0.45$ [Duchene \& Kraus, 2013]. The former is given by [Bovy 2017].

$d_{{\rm max},i}$ in Eq.~\ref{eq:number_systems} is the maximum distance 
corresponding to the given magnitude limit:
\begin{equation}
d_{{\rm max},i} = \left(\frac{L_{i}}{4 \pi F_{{\rm min}}}\right)^{1/2},
\label{eq:d_max}
\end{equation}
where the limiting flux in the bandpass $F_{{\rm min}}$ can also be 
stated in terms of the limiting magnitude $m_{{\rm min}}$,
\begin{equation}
m_{{\rm min}} = m_{0} - \frac{5}{2} \log_{10} \left(\frac{F_{{\rm 
min}}}{F_{0}}\right),
\end{equation}
for $m_{0}$ a zero-point magnitude and $F_{0}$ its corresponding flux (as 
always, everything is implicitly written in a defined bandpass).

In Eq.~\ref{eq:d_max}, again $i\in{\rm \{single,double\} }$, and as a 
consequence the maximum distance to which binary stars will be selected is 
greater than that of single stars, simply as a consequence of imposing a 
magnitude cut.
The ratio of double to single systems is
\begin{align}
\frac{N_d}{N_s} &= 
	\frac{n_d}{n_s} \left(\frac{d_{{\rm max}, d} }{d_{{\rm max}, s}}\right)^3 \\
&= {\rm BF} \times (1+\gamma_R)^{3/2}.
\end{align}
In the nominal case of twin binaries ($\gamma_R = 1$), with a binary fraction 
${\rm BF} = 0.5$, there are 
$\sqrt{2}$ more binary systems than single systems in the sample.
Correspondingly, there are $2\sqrt{2}$ more stars in binary systems than stars 
in single systems.

As a comment on Eq.~\ref{eq:number_systems}, if 
we wished to write a stellar number density profile that accounted for the 
vertical structure of the Milky Way, we might choose a profile either 
$\propto \exp(-z/H)$, or $\propto {\rm sech}^2(z/H)$ for $z$ the distance from 
the galactic midplane and $H$ a scale-height. Both density profiles would lead 
closed form analytic solutions.

\subsection{How many planets are in the sample?}

The number of planets in the sample is
\begin{align}
N_{\rm planets} =& N_{\rm planets\ in\ single\ star\ systems}  +  \\
				  &\quad\quad N_{\rm planets\ in\ double\ star\ systems} 
				  \nonumber \\
			   =& \Gamma_{t,s} N_s + 2 \Gamma_{t,d} N_d.
\end{align}

The factor of 2 accounts for the fact that there are twice as many stars in 
double star systems.

\subsection{What is the true occurrence rate?}
\label{sec:true_rate}

The ``true occurrence rate'' is the average number of planets per star. Thus

\begin{align}
\Gamma_t &= \frac{N_{\rm planets}}{N_{\rm stars}} \\
\Gamma_t &= \frac{\Gamma_{t,s} N_s + 2 \Gamma_{t,d} N_d}{N_s + 2N_d}.
\label{eq:true_occ}
\end{align}

\subsection{How many planets are detected?}
The total number of planet detections is the sum of the number of planets 
detected in single star systems $N_{{\rm det},s}$ and the number of planets 
detected in double star systems $N_{{\rm det},d}$.
These can be expressed individually. 
The former is
\begin{equation}
N_{{\rm det},s} = N_s \Gamma_{t,s} f_{s,{\rm geom}} f_{s,{\rm S/N} > {\rm 
(S/N)_{min}}},
\label{eq:N_det_s}
\end{equation}
where the product $N_s \Gamma_{t,s}$ is the number of planets in the single 
star systems of the sample, $f_{s,{\rm geom}}\equiv f_{s,g}$ is the geometric 
transit probability, and $f_{s,({\rm S/N} > {\rm (S/N)_{min}})} \equiv f_{s,c}$ 
is the fraction of these transiting planets that are observed with signal to 
noise greater than the minimum detection threshold (the completeness). 
Analogously,
\begin{equation}
N_{{\rm det},d} = 2 N_d \Gamma_{t,d} f_{d,g} f_{d,c},
\label{eq:N_det_d}
\end{equation}
where now $2 N_d \Gamma_{t,d}$ is the number of planets in the double star 
systems of the sample, the geometric transit probability is the same (if we 
have twin binaries -- not if we consider a distribution of secondaries) and the 
completeness term must account for any differences in the signal to noise 
distribution that come from stellar binarity.

\subsubsection{Analytic completeness}
Since the geometric transit probability is known, the only terms we have yet to 
compute are the completeness terms, $f_{i,c}$ for $i\in{\rm \{single,double\} 
}$. We proceed as follows.

The signal ${\rm S}$ for a box-car train transiting planet is
\begin{align}
{\rm S} &= \delta \mathcal{D} \\
&= \left(\frac{R_p}{R_\star}\right)^2 \mathcal{D},
\end{align}
for $R_p$ the planet's radius, $R_\star$ that of its host star, and 
$\mathcal{D}$ the dilution parameter defined as
\begin{align}
  \mathcal{D} &= 
  \Bigg\{\begin{array}{lr}
  L_1 / L_d, & {\rm\ if\ binary\ and\ target\ primary}\\
  \gamma_R L_1 / L_d, & {\rm\ if\ binary\ and\ target\ 
  secondary}\\
  1, & {\rm\ if\ single},
  \end{array} \nonumber\\
  &=
    \Bigg\{\begin{array}{lr}
    (1+\gamma_R)^{-1}, & {\rm\ if\ binary\ and\ target\ primary}\\
    (1 + \gamma_R^{-1})^{-1}, & {\rm\ if\ binary\ and\ target\ 
    	secondary}\\
    1, & {\rm\ if\ single},
    \end{array} 
  \label{eq:dilution}
\end{align}
where $L_1, L_d,$ and $\gamma_R$ were defined in the opening monograph.

Assuming the only source of noise is Poissonian counting noise, the noise ${\rm 
N}$ can be written
\begin{equation}
{\rm N} = \frac{1}{\sqrt{N_\gamma}},
\end{equation}
for $N_\gamma$ the number of photons received by the detector. This noise model 
is a useful simplification -- see [Howell 2006, pg 75] for the full 
CCD equation.
The number of received photons can be written
\begin{equation}
N_\gamma = F^{\rm N}_\gamma A N_{\rm tra} T_{\rm dur},
\end{equation}
for $F^{\rm N}_\gamma$ the photon number flux from the system [${\rm 
ph\,cm^{-2}\,s^{-1}}$], 
$A$ the detector area, $T_{\rm dur}$ the transit duration, and $N_{\rm tra} $ 
the number of transits observed, which is multiplied in assuming the transits 
are ``phase-folded''.

Thus the signal to noise ratio can be written
\begin{equation}
{\rm S/N} = \delta \mathcal{D} \sqrt{F^{\rm N}_\gamma A N_{\rm tra} T_{\rm 
dur}}.
\label{eq:snr_ivory_tower}
\end{equation}

This means we can write the minimum number flux of photons required for a 
detection at threshold as
\begin{equation}
F_{\rm lim}^N = \left[ \left({\rm \frac{S}{N}}\right)_{\rm min} \frac{1}{\delta 
	\mathcal{D}} \right]^2 \frac{1}{A N_{\rm tra} T_{\rm dur}}.
\end{equation}
To convert this to $F_{\rm lim}$, multiply by the average photon energy in the 
bandpass.

In passing, given the parameters that define a survey and planet type, 
Eq.~\ref{eq:snr_ivory_tower} would need to be re-expressed with 
$N_{\rm tra}$ roughly the ratio of the observing baseline to the planet period, 
and $T_{\rm dur}$ a function of $R_\star, P, a$, and impact parameter $b$, and 
then perhaps averaged over $b$. We leave them as-is for subsequent 
development.

The interesting term in Eq.~\ref{eq:snr_ivory_tower} that changes between stars 
of the same binarity class in our idealized sample is the square root of 
$F^{\rm N}_\gamma$.
This is the term that leads to a distribution of signal to noises for different 
stars.
The completenesses $f_{i,c}$ can be directly 
expressed in terms of those probability density functions:
\begin{equation}
f_{i,c} = 
	\int_{{\rm (S/N)_{min}}}^{\infty} 
		{\rm d}\left(\frac{{\rm S}}{{\rm N}}\right)_i \ 
		{\rm prob}\left(\frac{{\rm S}}{{\rm N}}\right)_i.
\label{eq:completeness_long}
\end{equation}
We keep the subscript $i$ because the signal to noise distributions are 
different for the cases of single star systems ($i=s$) and double star systems 
($i=d$).
Notably:
\begin{itemize}
	\item The dilution differs (Eq.~\ref{eq:dilution}).
	\item The photon number flux from the system differs.
\end{itemize}

To simplify notation, we let $x_i \equiv {\rm (S/N)}_i$, and rewrite 
Eq.~\ref{eq:completeness_long} as
\begin{equation}
f_{i,c} = 
\int_{{\rm x_{\rm min}}}^{\infty} 
{\rm d}x_i \ 
{\rm prob}(x_i).
\label{eq:completeness_short}
\end{equation}

\subsubsection{Deriving ${\rm prob}(x_i)$}
We want expressions for the probability density function of the observed signal 
to noise ratio, ${\rm prob}(x_i)$, for both the single and binary 
system case.

First, note that a star placed uniformly in the volume of the search space will 
have a probability density function for its distance $r$ from the origin of
\begin{equation}
{\rm prob} (r) = \frac{3 r^2}{d_{\rm max}^3},
\end{equation}
where the appropriate maximum distances should be substituted per 
Eq.~\ref{eq:d_max}.
Noting the transformation rule for probability density functions, we can 
evaluate the probability of a star having a observed number flux $F^{\rm 
N}_{i,\gamma}$ in the bandpass,
\begin{align}
{\rm prob} (F^{\rm N}_{i,\gamma}) 
&= {\rm prob}(r(F^{\rm N}_{i,\gamma}))
	\left| \frac{{\rm d} r}{{\rm d} F^{\rm N}_{i,\gamma}} \right| \\
&= \frac{3}{2 d_{\rm max}^3} c_i^{3/2} (F^{\rm N}_{i,\gamma})^{-5/2},
\label{eq:pdf_observed_flux}
\end{align}
where in the latter equality we have written a ``number luminosity'' $c_i$ 
(units of inverse time) defined for $i\in{\rm \{single,double\} }$ as
\begin{equation}
c_i = 
\Bigg\{\begin{array}{lr}
R_1^2 F^{\rm N}_{s1,\gamma}, & {\rm\ if\ single}\\
R_1^2 F^{\rm N}_{s1,\gamma} + R_2^2 F^{\rm N}_{s2,\gamma} & {\rm\ if\ double}.
\end{array}
\label{eq:c_i}
\end{equation}
In Eq.~\ref{eq:c_i}, $F^{\rm N}_{s1,\gamma}$ and $F^{\rm N}_{s2,\gamma}$ are 
the photon number fluxes at the surfaces of the stars. To derive 
Eq.~\ref{eq:pdf_observed_flux}, we simply scaled these by the distance:
\begin{equation}
F^{\rm N}_{i,\gamma} = \frac{c_i}{r^2}.
\label{eq:flux_at_detector}
\end{equation}

The surface photon number fluxes $F^{\rm N}_{si,\gamma}$ in Eq.~\ref{eq:c_i} 
are usually evaluated numerically, by convolving the wavelength-specific photon 
flux density of a star with the dimensionless spectral response function of the 
instrument. In other words,
\begin{equation}
F^{\rm N}_{s,\gamma} = \int F_\lambda T_\lambda \, {\rm d}\lambda.
\label{eq:pfd}
\end{equation}
The wavelength-specific photon flux density $F_\lambda$ [${\rm 
ph\,cm^{-2}\,s^{-1}\,\AA^{-1}}$] could be from Pickles' library, or could be 
a blackbody function. The transmission function is, up to factor of order 
unity, a step function between two wavelengths $\lambda_{\rm min}$ and 
$\lambda_{\rm max}$.
If we assume a blackbody source, Eq.~\ref{eq:pfd} becomes
\begin{equation}
F^{\rm N}_{s,\gamma} = 8\pi c \left(\frac{k T}{h c}\right)^3 
\int_{hc/(\lambda_{\rm max} kT)}^{hc/(\lambda_{\rm min} kT)}
\frac{u^2}{e^u - 1} \,{\rm d} u,
\end{equation}
which can be evaluated numerically\footnote{It may help in the numerics to note 
that infinite series representations of this type of dimensionless integral 
exist and converge quickly. For instance, one can show that
\begin{equation}
\int_{0}^{a} \frac{u^3}{e^u -1} \,{\rm d}u = \sum_{n=1}^{\infty} \frac{6}{n^4} 
- \frac{e^{-an}}{n^4} (6 + 6an + 3(an)^2 + (an)^3).
\nonumber
\end{equation}
A similar expression exists for the similar integral in the text. 
[Michels 1968] explains an analogous derivation.
}.

The importance of the functional form of $F^{\rm N}_{s,\gamma}$ is that it is 
to first order only a function of the blackbody temperature and the bandpass 
wavelength bounds.
Thus in the most general case $c_i(R_1, R_2, T_1, T_2, \lambda_{\rm min}, 
\lambda_{\rm max})$ and nothing else.
The only random variable involved in the flux being received at the detector is 
$r$, so we can indeed write the flux received at the detector as in 
Eq.~\ref{eq:flux_at_detector}.

We can finally write out the probability density functions for the signal to 
noise ratios in the single and double-star cases by using the transformation 
rule for pdfs, and applying Eq.~\ref{eq:snr_ivory_tower}.
For single stars,
\begin{equation}
{\rm prob}(x_s) = \frac{3}{d_{\rm max, s}^3} c_s^{3/2} \delta^3 \left( A T_{\rm 
dur} N_{\rm tra}\right)^{3/2} x_s^{-4}.
\label{eq:prob_xs}
\end{equation}
Analogously for double stars,
\begin{equation}
{\rm prob}(x_d) = \frac{3}{d_{\rm max, d}^3} c_d^{3/2} (\mathcal{D}\delta)^3 
\left( A T_{\rm dur} N_{\rm tra}\right)^{3/2} x_d^{-4}.
\label{eq:prob_xd}
\end{equation}

\subsubsection{Number of detected planets}

Performing the integrals of Eq.~\ref{eq:completeness_short}, we get:

\begin{align}
N_{{\rm det},s} &= N_s \Gamma_{t,s} f_{s,g} f_{s,c} \\
&= N_s \Gamma_{t,s} \frac{R_\star}{a} \frac{1}{d_{\rm max,s}^3} c_s^{3/2} 
\delta^3 \left(A 
T_{\rm dur} N_{\rm tra}\right)^{3/2} x_{\rm min}^{-3},
\end{align}
and
\begin{align}
N_{{\rm det},d} &= 2 N_d \Gamma_{t,d} f_{d,g} f_{d,c} \\
&= 2 N_d \Gamma_{t,d} \frac{R_\star}{a}
\frac{1}{d_{\rm max,d}^3} c_d^{3/2} (\mathcal{D}\delta)^3 \left(A T_{\rm dur} 
N_{\rm tra}\right)^{3/2} x_{\rm min}^{-3}.
\end{align}

Formally, the $f_{i,c}$ terms should be written as ${\rm min}(1, \ldots)$, 
where $(\ldots)$ is the given expression. This ensures that the fraction of 
planets above the signal to noise threshold is less than 1.
Note that we assumed twin binaries in the geometric transit probabilities.
Interestingly, the ratio of the two completeness fractions is
\begin{equation}
\frac{f_{d,c}}{f_{s,c}} = \left(\frac{d_{\rm max,s}}{d_{\rm max,d}}\right)^3 
\left(\frac{c_d}{c_s}\right)^{3/2} \mathcal{D}^3 = (1+\gamma_R)^{-3},
\end{equation}
which, for the $\gamma_R = 1$ case, gives 
$1/8$. (We got the same number from simple scaling laws at the outset, but the 
above machinery will be useful when generalizing in Sec.~\ref{sec:model_2}.)

The number of detected planets $N_{\rm det}$ is the sum of the two appropriate 
equations above, and can be written
\begin{align}
N_{\rm det} = 
&\left( \frac{\delta}{x_{\rm min}} \right)^3 (A T_{\rm dur} N_{\rm tra})^{3/2}
\frac{R_\star}{a}
\nonumber\\
&\quad \times
\left[ N_s \Gamma_{t,s} \frac{c_s^{3/2}}{d_{\rm max,s}^3}  +
	   2 N_d \Gamma_{t,d} \frac{c_d^{3/2}}{d_{\rm max,d}^3}  \right].
\label{eq:N_det}
\end{align}
All terms can be input to a computer, and checked against a Monte Carlo 
simulation if desired.

\subsection{Astronomer A ignores binarity}
Astronomer A has never heard of binary star systems. 
Nor has he heard of completeness corrections.
He knows about geometric transit probabilities.
What occurrence rate does he derive for planets of radius $R_p$ and period $P$?

The {\it total} occurrence rate (number of planets divided by number of 
``stars'') for Astronomer A would be $N_{\rm det}/(N_s + N_d)$, up to the 
geometric correction. 
However, even though Astronomer A does not know about binaries, the radii he 
derives for any planets in binary systems are too small, by a factor 
$\sqrt{\mathcal{D}}$ (for twin binaries, they are all $R_p/\sqrt{2}$).
Astronomer A wants an occurrence rate for planets of radius $R_p$ and period 
$P$.
The answer is thus
\begin{equation}
\Gamma_{{\rm A}, R_p} = \frac{N_{\rm det,s}/f_{s,g}}{N_s + N_d}.
\end{equation}

This astronomer will also think there is a second population of planets, with 
radius $R_p \sqrt{\mathcal{D}}$, and will thus rush to Nature claiming to also
have derived a second occurrence rate,
\begin{equation}
\Gamma_{{\rm A}, R_p \sqrt{\mathcal{D}}} = \frac{N_{\rm det,d}/f_{s,g}}{N_s 
+ N_d}.
\end{equation}
Note in the second case that this astronomer thinks these are single stars, and 
generally will miscompute the geometric transit probability.

\subsection{Astronomer B counts host stars correctly}
Astronomer B can somehow account correctly for the ``2 for 1'' 
effect of binarity, \textit{i.e.} that the sample actually has more stars than 
astronomer A thought.

By the same token as above,
\begin{equation}
\Gamma_{{\rm B},R_p} = \frac{N_{\rm det,s}/f_{s,g}}{N_s + 2N_d},
\end{equation}
and
\begin{equation}
\Gamma_{{\rm B},R_p \sqrt{\mathcal{D}}} = 
		\frac{N_{\rm det,d}/f_{d,g}}{N_s + 2N_d}.
\end{equation}

\subsection{Astronomer C counts host stars correctly and figures out diluted 
radii}
Astronomer C did high resolution imaging followup on every candidate, and 
correctly classifies the planetary radii.
Thus, she also knows which planets are in binary systems, and which are in 
single star systems.

She knows that the purported population of planets with radii $R_p 
\sqrt{\mathcal{D}}$ does not exist. All detected planets from this survey have 
radii $R_p$. She computes an occurrence rate
\begin{equation}
\Gamma_{{\rm C},R_p} = 
\frac{N_{\rm det,s}/f_{s,g} + N_{\rm det,d}/f_{d,g}}{N_s + 2N_d}
\end{equation}

the closest yet to the true rate (Sec.~\ref{sec:true_rate}).

\subsection{Astronomer D counts host stars correctly, figures out diluted 
radii, 
and accounts for completeness}

Astronomer D, knows which detections were around 
binaries and the associated radius correction.
They do injection recovery, and derive correct estimates for their 
completeness functions about single stars $f_{s,c}$, and 
about double star systems, $f_{d,c}$.
With this knowledge in hand, they compute the respective single and binary 
occurrence rates
\begin{equation}
\Gamma_{t,s} = \frac{N_{\rm det,s}}{N_s f_{s,g} f_{s,c}},
\end{equation}
\begin{equation}
\Gamma_{t,d} = \frac{N_{\rm det,d}}{2 N_d f_{d,g} f_{d,c}}.
\end{equation}

With these in hand, they derive the overall occurrence rate
\begin{align}
\Gamma_{{\rm D}, R_p} 
&= \frac{N_{\rm det,s}/(f_{s,g}f_{s,c}) + N_{\rm det,d}/(f_{d,g}f_{d,c})  }{N_s 
+ 2 N_d} \\
&= \frac{\Gamma_{t,s} N_s + 2\Gamma_{t,d} N_d}{N_s + 2 N_d}.
\end{align}

Although we admit it's been a bit of a slog, it happens that Astronomer D's 
occurrence rate is the true occurrence rate (cf. Eq.~\ref{eq:true_occ}).

All it takes is $\approx$ a full semester at Keck, a system-by-system analysis, 
and perfect understanding of the completeness of the detection efficiency for 
single and double star systems.

\subsection{Numerical verification}

To check the preceding analytic development, we implemented a Monte Carlo 
simulation of this idealized transit survey.
To run the survey, we defined the instrument specifications (detector area and 
transmission function), the stellar population (binary fraction, total number 
density of a given stellar class, fixed stellar 
properties), the planet population (fixed planet radius, period, and occurrence 
rate about single and binary stars), and finally the survey parameters 
(observing baseline, minimum SNR for ``detection'').
We then randomly drew star positions, randomly assigned planets to stars in 
single and binary systems, and computed the resulting signal to noise 
(Eq.~\ref{eq:snr_ivory_tower}) with which 
the transits would be observed.
As in the preceding analytics, we assumed ``twin'' binaries (same stellar 
radii, same effective 
temperature, and dilution does not depend on which stellar binary is the 
``target'').

The results are shown in Fig.~\ref{fig:snr_dist_analytic_v_numeric}, and 
indicate that the analytic probability distribution functions 
Eqs.~\ref{eq:prob_xs},~\ref{eq:prob_xd}, and the number of 
detections (Eq.~\ref{eq:N_det}) are correct.

A point evident in Fig.~\ref{fig:snr_dist_analytic_v_numeric} is that, for 
fixed planet parameters, and fixed stellar parameters ($R_\star, L_\star$, and 
distance $r$) the SNR distribution for planets in binaries is poorer than that 
of planets in single star systems.
We can see analytically that this simply due to dilution:
\begin{equation}
\frac{{\rm prob}(x_d)}{{\rm prob}(x_s)} =(1 + \gamma_R)^{-1} = \mathcal{D}.
\end{equation}
Deriving this simple form requires noting that the ratios of the 
bandpass-specific number luminosities (Eq.~\ref{eq:c_i}) is equal to the ratio 
of the bandpass-specific energy luminosities (otherwise a term with $c_s/c_d$ 
must be included).

\begin{figure}[!t]
	\begin{center}
		\includegraphics[scale=0.4]{figures/snr_distribution.pdf}
	\end{center}
	\caption{Comparison of analytic and numeric probability 
	density functions of the SNR in an idealized transit survey.
	The analytic lines are Eqs.~\ref{eq:prob_xs} and~\ref{eq:prob_xd} for the 
	planet populations orbiting single and binary stars. The underlying stepped 
	histogram is output from Monte Carlo simulations. Poisson noise leads to 
	a small deviation at the faint and bright limits, but the numerics and 
	analytics otherwise agree.
		 }
	\label{fig:snr_dist_analytic_v_numeric}
\end{figure}

\subsection{Representative numbers for a few cases}

\subsubsection{Twin binaries: if we ignore binarity, for what fraction of 
detections do we misclassify the radii?}

Ignoring binarity, we will detect $N_{\rm det,s}$ planets around single stars, 
and $N_{\rm det,d}$ planets around double stars. The latter set will be assumed 
to have radii $R_p \sqrt{\mathcal{D}}$.
The fraction of detections with misclassified radii can then be written
\begin{equation}
\frac{N_{\rm det,d}}{N_{\rm det,s} + N_{\rm det,d}} = \frac{1}{1+\alpha},
\end{equation}
for
\begin{equation}
\alpha \equiv \frac{1}{2({\rm BF})} (1+\gamma_R)^{3/2} 
\frac{\Gamma_{t,s}}{\Gamma_{t,d}}.
\end{equation}

For the nominal G2V dwarf case of ${\rm BF}=0.45$, twin binaries with equal 
occurrence rates this produces a misclassification rate of $24\%$, in agreement 
with Fig.~\ref{fig:snr_dist_analytic_v_numeric}.

\subsubsection{Twin binaries: if we ignore binarity, how wrong is our
occurrence rate for planets of radius $R_p$?}

This is almost simply asking ``what is the relative difference between the 
occurrence rates derived by Astronomers D and A for planets of radius $R_p$?''
However, in the more realistic case, Astronomer A also has derived a 
completeness, which we assume is the same as for Astronomer D in the single 
star case.
So Astronomer A now misclassifies planetary radii, and miscounts the total 
number of stars, but knows his completeness for single stars.
Astronomer D corrects all these errors.
Then
\begin{equation}
\Gamma_{{\rm A}, R_p} = \frac{N_{\rm det,s}/(f_{s,g} f_{s,c})}{N_s + N_d},
\end{equation}
and the relative difference between the two occurrence rates is

\begin{align}
\left|\frac{\Gamma_{{\rm A},R_p} - \Gamma_{{\rm D},R_p}}
		   {\Gamma_{{\rm A},R_p}} \right| &=
\left| 1 - \frac{\Gamma_{{\rm D},R_p}}{\Gamma_{{\rm A},R_p}}\right| \\
&=
\left| 1 - \left(\frac{\Gamma_{t,s}N_s + 2\Gamma_{t,d}N_d}{N_s + 2N_d}  \cdot 
\frac{N_s + N_d}{\Gamma_{t,s} N_s }\right)\right| \\
&= 
\left|
1 - \frac{(1 + 2 \beta \Gamma_{t,d}/\Gamma_{t,s})(1 + \beta)}{(1+2\beta)}
\right|,
\end{align}
for
\begin{equation}
\beta \equiv N_d/N_s = {\rm BF} \times (1+\gamma_R)^{3/2}.
\end{equation}

For the nominal G2V dwarf case of ${\rm BF}=0.45$ with twin binaries ($\gamma_R 
= 1$) and $\Gamma_{t,d} = \Gamma_{t,s}$ this gives a relative error of $127\%$.
For instance, in the numerical simulation corresponding to 
Fig.~\ref{fig:snr_dist_analytic_v_numeric}, Astronomer A finds $\Gamma_{\rm 
A,R_p} = 0.22$, while Astronomer 
D derives the true (input) occurrence rate of $\Gamma_{\rm D,R_p} = 0.5$.

\section{Model \#2: fixed stars, fixed planets, varying light-ratio binaries}
\label{sec:model_2}

>>>>

\documentclass[12pt,modern]{aastex61}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{todonotes}

@@

\title{ Binarity's effects on occurrence rates measured by transit surveys}

@@

\input{generalities}
\input{model_1}
\input{model_2}
